Statistical Analysis of Clinical Phenotype and Patient Survival in Neurodevelopmental Disorder with Microcephaly, Arthrogryposis, and Structural Brain Anomalies

### 2022-09-07 paper/paper.qmd
updated template

---
title: "Statistical Analysis of Clinical Phenotype and Patient Survival in Neurodevelopmental Disorder with Microcephaly, Arthrogryposis, and Structural Brain Anomalies"
author: "Dean Marchiori"
format: pdf
abstract: "A recently described, rare genetic condition known as Neurodevelopmental Disorder with Microcephaly, Arthrogryposis, and Structural Brain Anomalies (NEDMABA) has been identified in children with bi-allelic loss-of-function variants in *SMPD4*. The progression of this condition is not well understood with the limited case reports described so far exhibiting a severe and clinically diverse phenotype. A gap exists in the understanding the associations of the heterogenous features present in the clinical phenotype and the expected survival probabilities of affected individuals. This is driven in part to the paucity of analysis-ready data in reported cases. This research aims to collate and standardise available case reports to analyse and identify meaningful clusters in the clinical phenotype, and to quantify the survival probability for children with NEDMABA. To overcome the challenge of sparse, multidimensional data on very few subjects, we employ Multiple Correspondence Analysis (MCA) as a dimension reduction technique, which is then subject to cluster analysis and interpretation. To quantify more accurate survival probabilities, Kaplan-Meier estimation is formulated to account for censoring in the data. The analysis identified the classic phenotype for this condition, as well as one other distinct feature cluster relating to findings of vocal cord paralysis and respiratory failure. The survival probability for those affected was found to decline sharply in early infancy with median survival of 150 days, but with some as high as 12.5 years. This wide range of outcomes is provisionally associated with variant type however interpretation of these results are guarded based on very low sample sizes. Despite challenges with sparse and inconsistent data, this analysis represents the first of its kind to help describe associations and trajectories of individuals with this condition and can provide clinicians and genetic counsellors with better information to aide in decision making and support for families."
editor: visual
bibliography: refs.bib
execute: 
  echo: false 
---

@@

```{r}
# table with med gen 
```

@@

# References

>>>>

---
title: Statistical Analysis of Clinical Phenotype and Patient Survival in Neurodevelopmental Disorder with Microcephaly, Arthrogryposis, and Structural Brain Anomalies
author:
  - name: Dean Marchiori
    email: deanmarchiori@gmail.com
    attributes:
        corresponding: true
abstract: |
  A recently described, rare genetic condition known as Neurodevelopmental Disorder with Microcephaly, Arthrogryposis, and Structural Brain Anomalies (NEDMABA) has been identified in children with bi-allelic loss-of-function variants in *SMPD4*. The progression of this condition is not well understood with the limited case reports described so far exhibiting a severe and clinically diverse phenotype. A gap exists in the understanding the associations of the heterogenous features present in the clinical phenotype and the expected survival probabilities of affected individuals. This is driven in part to the paucity of analysis-ready data in reported cases. This research aims to collate and standardise available case reports to analyse and identify meaningful clusters in the clinical phenotype, and to quantify the survival probability for children with NEDMABA. To overcome the challenge of sparse, multidimensional data on very few subjects, we employ Multiple Correspondence Analysis (MCA) as a dimension reduction technique, which is then subject to cluster analysis and interpretation. To quantify more accurate survival probabilities, Kaplan-Meier estimation is formulated to account for censoring in the data. The analysis identified the classic phenotype for this condition, as well as one other distinct feature cluster relating to findings of vocal cord paralysis and respiratory failure. The survival probability for those affected was found to decline sharply in early infancy with median survival of 150 days, but with some as high as 12.5 years. This wide range of outcomes is provisionally associated with variant type however interpretation of these results are guarded based on very low sample sizes. Despite challenges with sparse and inconsistent data, this analysis represents the first of its kind to help describe associations and trajectories of individuals with this condition and can provide clinicians and genetic counsellors with better information to aide in decision making and support for families.
keywords: 
  - NEDMABA
  - SMPD4
#journal: "An awesome journal"
date: last-modified
execute: 
  echo: false 
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

@@

# References {-}
